Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-07-27 surfaces-write-up/surface-write-up.tex
surfaces: comment on todos in hirz, ker spacing

By Lemma ~\ref{lem:bound-ker-chi}, we can bound the degree of generation of
$\ker(\chi)$ and we next calculate the degree of generation of $\ker \psi$:

@@

By ~\ref{lem:bound-ker-chi}, $\ker(\chi)$ is generated in degrees 
up to $2\sum_{i=1}^n \ell_i a_i$ and by Lemma
~\ref{lem:proj-relations-psi}, $\ker \psi,$ is generated in
degrees up to $\frac{\bida_i}{\deg(D)} + \sum_{i=1}^n \ell_i a_i$. 
Consequently, Lemma ~\ref{lem:composite-map} implies that $\ker \phi$
is generated in degrees less than
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}
	(\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]
\end{proof}

@@

Aaron: I did
a search for $\epsilon$ and the last time it is mentioned is inside this proof.
In fact, while the $\epsilon$ are not referred to later, the $e_{ij}$ are referred to!
The ordered tuple will look slightly longer, but it is less confusing then
switching from tuple notation to
product notation for no reason, especially when it isn't used again.
It's confusing for the reader who is thinking in terms of lattice points
to suddently switch to thinking in a ring. At the very least, you should say 
it is the preimage of the map from the lattice to the canonical ring, since the
current use of corresponding isn't ever made clear, but
this would probably then take a bunch more space too, and I would also find this significantly more confusing than just saying the lattice points.}
A quick calculation deduces that these $\epsilon_i$'s and $\epsilon_{i,j}$'s lie on the extremal rays represented by the $e_i$'s and $e_{i,j}$'s, respectively.
\begin{rem}
For the remainder of this section replace the $e_i$'s and $e_{i,j}$'s with a (possibly) different representative of it's extremal ray say that their lattice points correspond to the $\epsilon_i$'s and $\epsilon_{i,j}$'s, respectively.
\end{rem}

\todo{Peter: I don't like having a remark inside of a proof.  Also, it feels
like this could reasonably be replaced with a better constructed/organized
proof, so that $e_i$'s are just defined this way initially; let me know if you
see a good way to do so...  As for Aaron's comment above, we could do something
like that, except we already defined $e_i$'s to be something different, which
might be equally/more confusing.
Aaron: I also don't like this here, just because it isn't really a remark,
its more of a notational convention, and it doesn't really belong in the proof.
I also think it's confusing to switch notation in the middle of a proof.
Earlier in the proof, we should just call the $e_{ij}$ something like $\hat{e_{ij}}$. Then, we should never define $\epsilon_{ij}$, since it's only used in this
proof, and should really be $e_{ij}$, as I was explaining in my previous todo.
Instead of defining $\epsilon_{ij}$, just define $e_{ij}$ which are the
appropriately scaled $\hat{e_{ij}}$ and everything will be much clearer.}
Thus,
Proposition ~\ref{prop:cone-generation} implies that 
$R_D$ is generated in 
degrees less than the sum of the degrees of the $\epsilon_i$'s and $\epsilon_{i,j}$'s, which is
\[
	\rho = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i + \sum_{\substack{
	i \in \Tp(D) \\	j \in \Tm(D)}} (\bida_i \bidb_j- \bida_j \bidb_i)\ell_{i,j}.
\]
\end{proof}

@@

By Lemma ~\ref{lem:bound-ker-chi}, we can bound the degree of generation of $\ker(\chi)$ below
$2 \rho$.
Finally, we calculate the degree of generation of $\psi$:

@@

\begin{proof}
The generation degree bound is as stated in Lemma \ref{lem:hirz-generators}.
By Proposition \ref{lem:composite-map}, the degree of generation of
$\ker \phi$ is at most the maximum of the generation degrees of $\ker(\chi)$
and $\ker \psi$, giving us the desired relations bound. The bound on $\ker(\chi)$
follows from Lemma ~\ref{lem:bound-ker-chi} and the bound on $\ker \psi$
follows from Lemma ~\ref{lem:hirz-bound-ker-psi}.
$\ker \psi$ follows from 
\end{proof}

>>>>

By Lemma ~\ref{lem:bound-ker-chi}, we can bound the degree of generation of
$\ker \chi $ and we next calculate the degree of generation of $\ker \psi$:

@@

By ~\ref{lem:bound-ker-chi}, $\ker \chi$ is generated in degrees 
up to $2\sum_{i=1}^n \ell_i a_i$ and by Lemma
~\ref{lem:proj-relations-psi}, $\ker \psi,$ is generated in
degrees up to $\frac{\bida_i}{\deg(D)} + \sum_{i=1}^n \ell_i a_i$. 
Consequently, Lemma ~\ref{lem:composite-map} implies that $\ker \phi$
is generated in degrees less than
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}
	(\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]
\end{proof}

@@

Aaron: I did
a search for $\epsilon$ and the last time it is mentioned is inside this proof.
In fact, while the $\epsilon$ are not referred to later, the $e_{ij}$ are referred to!
The ordered tuple will look slightly longer, but it is less confusing then
switching from tuple notation to
product notation for no reason, especially when it isn't used again.
It's confusing for the reader who is thinking in terms of lattice points
to suddently switch to thinking in a ring. At the very least, you should say 
it is the preimage of the map from the lattice to the canonical ring, since the
current use of corresponding isn't ever made clear, but
this would probably then take a bunch more space too, and I would also find this significantly more confusing than just saying the lattice points.

Robin: Yeah, I thought that we had agreed upon doing the entire
proof in terms of the lattice points and then occasionally giving
the algebraic translation (and noting that it is simply the
algebraic translation) rather than this ``in-between'' sort of
thing. I agree with Aaron that it's confusing for a reader who has
not gone through the details before.}

A quick calculation deduces that these $\epsilon_i$'s and
$\epsilon_{i,j}$'s lie on the extremal rays represented by the
$e_i$'s and $e_{i,j}$'s, respectively.
\begin{rem}
For the remainder of this section replace the $e_i$'s and $e_{i,j}$'s with a (possibly) different representative of it's extremal ray say that their lattice points correspond to the $\epsilon_i$'s and $\epsilon_{i,j}$'s, respectively.
\end{rem}

\todo{Peter: I don't like having a remark inside of a proof.  Also, it feels
like this could reasonably be replaced with a better constructed/organized
proof, so that $e_i$'s are just defined this way initially; let me know if you
see a good way to do so...  As for Aaron's comment above, we could do something
like that, except we already defined $e_i$'s to be something different, which
might be equally/more confusing.
Aaron: I also don't like this here, just because it isn't really a remark,
its more of a notational convention, and it doesn't really belong in the proof.
I also think it's confusing to switch notation in the middle of a proof.
Earlier in the proof, we should just call the $e_{ij}$ something like $\hat{e_{ij}}$. Then, we should never define $\epsilon_{ij}$, since it's only used in this
proof, and should really be $e_{ij}$, as I was explaining in my previous todo.
Instead of defining $\epsilon_{ij}$, just define $e_{ij}$ which are the
appropriately scaled $\hat{e_{ij}}$ and everything will be much clearer.
Robin: agree with Aaron here.}

Thus, Proposition ~\ref{prop:cone-generation} implies that 
$R_D$ is generated in 
degrees less than the sum of the degrees of the $\epsilon_i$'s and $\epsilon_{i,j}$'s, which is
\[
	\rho = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i + \sum_{\substack{
	i \in \Tp(D) \\	j \in \Tm(D)}} (\bida_i \bidb_j- \bida_j \bidb_i)\ell_{i,j}.
\]
\end{proof}

@@

By Lemma ~\ref{lem:bound-ker-chi}, we can bound the degree of generation of $\ker \chi$ below
$2 \rho$.
Finally, we calculate the degree of generation of $\psi$:

@@

\begin{proof}
The generation degree bound is as stated in Lemma \ref{lem:hirz-generators}.
By Proposition \ref{lem:composite-map}, the degree of generation of
$\ker \phi$ is at most the maximum of the generation degrees of $\ker \chi$
and $\ker \psi$, giving us the desired relations bound. The bound on $\ker \chi$
follows from Lemma ~\ref{lem:bound-ker-chi} and the bound on $\ker \psi$
follows from Lemma ~\ref{lem:hirz-bound-ker-psi}.
$\ker \psi$ follows from 
\end{proof}
